Moment Constraints Make Learning Latent-Variable Models Easier

### 2014-01-30 doc/icml2014/undirected.tex
Updated undirected section

In this section, we will show that we can extend the results to undirected graphical models.

Let $z$ be a collection of variables indexed by $V$ (that is, $z = \{z_j\}_{j \in V}$).
Suppose that some of the variables are observed ($x = \{ z_j \}_{j \in O}$, where $O \subset V$) 
and the rest are latent ($h = \{ z_j \}_{j \in H}$, where $H = V \backslash O$).
Let $C \subset 2^V$ be a collection of subsets of variables,
and let $\phi_c(z_c) \in \Re^d$ be the feature map for each $c \in C$.
In this paper, we will consider both directed and undirected models over $z$.
In the undirected case, we define the standard globally-normalized exponential family model:
\begin{align}
  \label{eqn:undirectedSetup}
  p_\theta(z) = \exp\{ \phi(z)^\top\theta - A(\theta) \},
\end{align}
where $\phi(z) \eqdef \sum_{c \in C} \phi_c(z_c) \in \Re^d$ is the global feature vector, $\theta \in \Re^d$ is the parameter vector,
and $A(\theta) = \log \int \exp\{\phi(z)^\top\theta\} dz$ is the global log-partition function.
To some extent, our methods only depend on the conditional independence
structure of the models.

\paragraph{Moment Constraints}

In the previous section, we saw that we could learn moments $P(x1,x2)$ given conditions. This makes the problem convex, boom.

\paragraph{Examples}

\paragraph{General latent-variable models with bottlenecks}

Our learning algorithm for three-view mixture models relies on the fact that
the sufficient statistics of the exponential family
can be expressed as a function of the latent moments.
Of course, these conditions do not hold for general log-linear models.
Rather than falling back on plain local optimization of the likelihood in these
cases, we would still like
to somehow use the \emph{partial information} offered by the latent moments.

Partial information is available when the following holds:
%\begin{property}
%  There exists some latent variable $h_{j_0} \in [\nh]$ (the bottleneck)
%  and three conditionally independent observed variables $x_{j_1},x_{j_2},x_{j_3}$ (views) such that
%  each conditional mean matrix $\E[\phix(x_{j}) \mid h_{j_0}] \in \R^{\nphix \times \nh}$ ($j \in \{ j_1, j_2, j_3 \}$) has full column rank $\nh$
%  for some transformation $\phix$.
%\end{property}

Optimization.

>>>>

In this section, we will show that we can extend the results to
  undirected graphical models parameterized as log-linear latent variable
  models.
We exploit a connection between the moments of the clique potentials and
  the log-partition function to show that learning parameters is convex
  given the clique marginals $Z_\sC$. 
The end result is a consistent estimator for log-linear
  latent-variable models which satisfy \propertyref{exclusive-views}.
Finally, we generalize the result to classes when only a subset of the
  clique marginals can be learned.

As before, let $\sG$ be a graphical model with variables $V$, with $X
  \subseteq V$ being observed and $H = V \setminus X$ hidden.
We assume that the undirected graphical model is a log-linear latent
  variable model with parameters $\theta$,
\begin{align*}
  p_\theta(\vx, \vh) &= \exp( \theta^T\phi(x,z) - A(\theta) ),
\end{align*}
where $\phi(x,z)$ is a feature vector with components $\phi_\sC(x,z)$
  for every clique and $A(\theta) = \sum_{\vx, \vh}  \exp(
  \theta^T\phi(x,z) )$ is the log-partition function.
While $\phi_\sC(x,z)$ can be an arbitrary featurization, we assume
  without loss of generality that $\phi_\sC(x,z)$ are the indicator
  features $\BI[x,z]$\footnote{Note that $\BI[x,z]$ are sufficient
  statistics for the model, and hence we do not lose any statistical
  efficiency in using them}.

As before, our objective is to find a set of parameters $\hat \theta$
  that maximize the likelihood of data $\sD$.
\begin{align}
  \hat \theta 
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log p_\theta(\vx) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} p_\theta(\vx,\vh) \nonumber \\
      &= \arg\max_{\theta} \sum_{\vx \in \sD} \log \sum_{\vh} \exp( \theta^T\phi(\vx,\vh) - A(\theta) ). \label{eqn:obj-ml}
\end{align}
Note that this problem is not concave in $\theta$.

An alternate objective is to optimize the following lower-bound on the log-likelihood, 
\begin{align}
  \sL_p(q,\theta) &= \sum_{\vx \in \sD} \sum_{\vh} q(\vh \given \vx) \log\frac{p_\theta(\vx,\vh)}{q(\vh \given \vx)} \\
    &= \KL( q(\vh \given \vx) || p_{\theta}(\vh \given \vx)) \label{eqn:obj-kl}
\end{align}

Importantly, given $q$, the above objective is convex in $\theta$ and vice versa.
It can be shown\cite{} that the local-optima of \equationref{obj-kl}
are also local-optima of \equationref{obj-ml}. Usually, one
iteratively optimizes $q$ and $p_\theta$ from some initial parameters
$\theta_0$ to arrive at a local optimum.

This reduces to,
\begin{align}
  \sL_p(\theta) &= \theta^T Z - A(\theta),
\end{align}
which is a convex problem.

However, from the previous section, for any model satisfying
  \property{exclusive-views}, we have a consistent estimator for the
  marginals $q \equiv Z$. 
Therefore, we can optimize for $\theta$ given this estimate of $q$,
  $\hat q$.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnLogLinear}
  \label{algo:undirected}
  \begin{algorithmic}
    \REQUIRE A graphical model $\sG$ satisfying \textbf{(P1)}, \textbf{(P2)}, data $\sD$
    \ENSURE A consistent estimate $\hat \theta$ of the parameters of the model.
      \STATE Apply \LearnMarginals to $\sG$ to learn clique marginals $Z_\sC$.
      \STATE Optimize \equationref{obj-kl} with respect to $\theta$.
  \end{algorithmic}
\end{algorithm}

This approach is outlined in \algorithmref{undirected}.

\todo{Do we have to worry about the fact that this is constrained?}
\todo{What about model un-identifiability?}

\paragraph{Learning with partial moment constraints}

In more general models, \propertyref{exclusive-views} may not hold for
  every clique in the graph, however we can still estimate the marginals
  $Z_\sC$ for cliques in which it does.
In this section, we will describe a natural method to use these
  constraints to simplify the parameter search problem.

Constrain the KL to be in $Q$.

\subsection{Examples}

\paragraph{Latent MRFs}

Image segmentation

Separate features into views

\todo{Some challenge example requiring measurements em}
